\chapter{Conference talks}
\label{app-e}

%\textbf{Attribution}

%I presented the following conference talks about Tangible Landscape:

Brendan A. Harmon, Helena Mitasova, and Anna Petrasova. 2014. Tangible geospatial modeling for landscape architects. In \emph{2014 Geodesign Summit}. Redlands, California: Esri.
\url{http://video.esri.com/watch/3170/tangible-geospatial-modeling-for-landscape-architects}

Brendan Harmon, Anna Petrasova, and Vaclav Petras. 2016. Serious gaming with Tangible Landscape. In \emph{Coffee \& Viz}. Raleigh, NC.
\url{https://baharmon.github.io/tei-2016/}

Brendan A. Harmon. 2016. Embodied spatial thinking in tangible computing. In \emph{TEI ’16: Tangible, Embedded, and Embodied Interaction 2016}. Eindhoven, Netherlands.
\url{http://ncsu-geoforall-lab.github.io/coffee-and-viz/hunt.html\#/9}

Brendan A. Harmon, Anna Petrasova, Vaclav Petras, Helena Mitasova, and Ross K. Meentemeyer. 2016. Creative spatial thinking with Tangible Landscape. In \emph{American Association of Geographers Annual Meeting 2016}. San Francisco, CA.
\url{http://baharmon.github.io/aag-2016/}

Brendan A. Harmon, Anna Petrasova, Vaclav Petras, Helena Mitasova, and Ross K. Meentemeyer. 2016. Tangible interaction for GIS. In \emph{FOSS4G NA 2016}. Raleigh, NC.
\url{http://baharmon.github.io/foss4g-na-2016/}

Brendan A. Harmon, Anna Petrasova, Vaclav Petras, Helena Mitasova, and Ross K. Meentemeyer. 2016. Tangible Landscape: cognitively grasping the flow of water. In \emph{International Society of Photogrammetry and Remote Sensing 2016}. Prague.
\url{https://baharmon.github.io/tangible-water-flow-presentation/}

Brendan A. Harmon, Anna Petrasova, Vaclav Petras, Helena Mitasova, and Ross K. Meentemeyer. 2016. Tangible geographies. In \emph{Royal Geographical Society Annual International Conference 2016}. London.
\url{https://baharmon.github.io/rgs-2016/}

% co-authored
%
%GIS-based modeling with tangible interaction
%Free and Open Source Software for Geospatial (FOSS4G) 2014
%\cite{Petrasova2014a}
%
%Using GRASS GIS through Python and tangible interfaces
%FOSS4G NA 2016
%\cite{Petrasova2016a}
%


\vfil
\pagebreak



